---
title: "Class 12: Structural Bioinformatics Part I"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Here we inspect the types of structures in the main database for biomolecular data - the PDB.

### Questions:

**Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy.**
```{r}
# Read in .csv file  
stats <- read.csv("Data Export Summary.csv", row.names=1)

# Calculate percentage
ans <- stats$Total/sum(stats$Total) * 100
names(ans) <- rownames(stats)
round(ans, 2)

```
**-- Also can you determine what proportion of structures are protein?**
```{r}
# Calculate percentage of proteins
proteins <- sum(stats$Proteins)/sum(stats$Total) * 100
round(proteins, 2)
```
**-- Aim to have a rendered GitHub document with working code that yields your answers.**

**Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB.**
-- There are currently 1289 protease structures in the current PDB.

