---
title: "Class 5: Data Visualization and Graphs in R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Visualization and graphs in R
```{r}
plot(1:10, col="blue", typ="o")
```

# Section 2: Customizing plots

Need to import/read input data file first
```{r}
baby <- read.table("bimm143_05_rstats/weight_chart.txt", header = TRUE)
```

A basic plot of age vs weight
```{r}
plot(baby$Age, baby$Weight, type = "o", col="pink",
    pch = 15, cex = 1.5, lwd = 2, ylim = c(2,10),
    xlab = "Age (months)", ylab = "Weight (kg)",
    main = "Baby Weight with Age ")
```

Import features_count.txt file
```{r}
mouse <- read.table("bimm143_05_rstats/feature_counts.txt", header = TRUE, sep = "\t")
```

Plot mouse data (counts only) into barplot
```{r}
par(mar = c(3, 12, 4, 2), las=1) #mar sets margins to fit plot, las sets ylabels horizontal
barplot(mouse$Count, horiz=TRUE, names.arg = mouse$Feature, xlim = c(0, 80000), main = "Number of Features in the Mouse GRCm38 Genome")
```

# Section 3: Using color in plots
Import file to plot later
```{r}
mf <- read.delim("bimm143_05_rstats/male_female_counts.txt")
```

Plot data into barplot with all bars different colors using rainbow()
```{r}
par(mar=c(7,4,2,2))
barplot(mf$Count, las=2, ylab = "Counts", names.arg = mf$Sample, ylim = c(0,20), col= rainbow(nrow(mf)))
```

Plot data with alternating colors for male/female
```{r}
barplot(mf$Count, las=2, ylab = "Counts", names.arg = mf$Sample, ylim = c(0,20), col= c("lightblue","pink"))
```
