---
title: "Class 18: Investigating Cancer Genomics Datasets"
author: "Larissa Rapadas"
date: "3/5/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The GenomicDataCommons Packages

First we install the three packages from Bioconductor: `GenomicDataCommons`, `TCGAbiolinks` and `maftools`. This is done in the console:
```{r}
#BiocManager::install( c("GenomicDataCommons", "TCGAbiolinks", "maftools") )

```

Now we have to load the newly downloaded packages into our console:
```{r}
library(GenomicDataCommons) # to talk to the database and download data all from R

library(TCGAbiolinks)
library(maftools)
```

Now let's check on the GDC status:
```{r}
status()
```

## Querying the GDC from R
```{r}
cases_by_project <- cases() %>%
  facet("project.project_id") %>%
  aggregations()
head(cases_by_project)

# pipe operator %>% -> organizes the sequence of functions
```

### Plot this overview data with a barplot:
```{r}
x <- cases_by_project$project.project_id
```
```{r}
barplot(x$doc_count, names.arg= x$key, log ="y", las=2, col="lightblue")
```

## Variant Analysis with R

Find files available for PAAD project that we can then examine variants from
```{r}
maf.file <- GDCquery_Maf(tumor="PAAD", pipelines = "mutect")

```
```{r}
head(maf.file)

```

Read in the MAF file we found:
```{r}
vars = read.maf(maf = maf.file, verbose = FALSE)

```

Now we make a summary plot of this data:
```{r}
plotmafSummary(vars)

```

## Oncoplots

A very useful summary representation of this data can be obtained via so-called oncoplots, also known as waterfall plots.

```{r}
oncoplot(maf = vars, top = 10)

```
```{r}

```

# Part II
Example analysis available on:
https://bioboot.github.io/bimm143_W20//class-material/lecture18_part2_example/

