---
title: "Class 13: Structural Bioninformatics Part II"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Download and clean HIV-Pr receptor structure

Here we will download the 1HSG PDB structure from the online database, remove the bound drug from Merck and write out a protein-only and ligand-only new set of PDB files

We will use the bio3d package functions `get.pdb()`, `read.pdb()`, `atom.select()`, and finally `write.pdb()`


```{r}
# Download the PDB structure
library (bio3d)
file.name <- get.pdb("1hsg") 

# Read the PDB structure
hiv <- read.pdb(file.name) 
hiv
```
### Questions

**Q1: What is the name of the two non protein** ***resid*** **values in this structure?**
-- HOH and MK1

**What does** ***resid*** **correspond to and how would you get a listing of all resid values in this structure?**
-- reside corresponds to the residue name in the protein

```{r}
prot <- atom.select (hiv, "protein", value=TRUE)
lig <- atom.select (hiv, "ligand", value=TRUE)
```

Select protein only
```{r}
write.pdb (prot, file="1hsg_protein.pdb")
```

Select ligand only
```{r}
write.pdb (lig, file="1hsg_ligand.pdb")
```

