---
title: 'Class 6: R Functions'
author: "Larissa Rapadas"
date: "1/23/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#this is a silly plot
plot (1:5, col= "purple", typ="o")
```

Let's see more about **file import** (i.e. reading files into R). The main read function in base R is `read.table()`

```{r}
read.csv("test1.txt") #by default has comma-separated values and headers
read.csv("test2.txt", sep="$") #has headers but is separated by "$"
read.table("test3.txt") #by default has space-separated values and NO headers
```

# Functions
Our first example function:
```{r}
add <- function(x, y=1) {
  #Sum the input of x and y
  x + y
}
```

Let's try using this function
```{r}
add(1)
add(7,3)
```

Let's see how this works with vectors as input:
```{r}
add(c(1,2,3)) #adds 1 to each element to vector c
add(c(1,2,3), 4) #adds 4 to each element to vector c
add(c(1,2,3), c(1,2,3)) #adds each element to the corresponding vector elements
```

What is this `range()` function you speak of?
```{r}
x <- c(4,4,10,3,11)
max(x)
min(x)
```
```{r}
range(x)
```

# This is our second function:
You need a “name”, “arguments” and “body”…
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

## Test on a small example where you know the answer
```{r}
rescale(1:10)
```

## How would you get your function to work here…
```{r}
rescale(c(1,2,NA,3,10))
```
Because we haven't set our function to omit NA's, we must argue with the function to do so
```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm=TRUE)
 (x - rng[1]) / (rng[2] - rng[1])
}
rescale2(x)
```
Now that we have done so, our previous function call should now work, omitting our NA input
```{r}
x <- rescale2(c(1,2,NA,3,10))
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 return(answer)
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```
```{r}
rescale3(x)
```
#Hands-on Section B

```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

**Q1.** What type of object is returned from the `read.pdb()` function?
- Returns a large list of 8 things and of class "pdb" and "sse"
```{r}
class(s1)
str(s1)
```
**Q2.** What does the trim.pdb() function do?
- Trims the previously read pdb file. Creates a new PDB object based on a subset of atoms

**Q3.** What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?
- 
**Q4.** What would be a better plot to compare across the different proteins?
- Combine the three separate plots into one for easier comparison of lines
```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, typ="l", col="blue", lwd=1)
points(s3.b, typ="l", col="red", lwd=1) 
```

**Q5.** Which proteins are more similar to each other in their B-factor trends. How could you quantify this?
- With a cluster dendrogram. This quantifies the distance between the the rows of the three different vectors below.
```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

**Q6.** How would you generalize the original code above to work with any set of input protein structures?
c(read.pdb("4AKE"),)
