---
title: "HW_Q6: Creating R Functions"
author: "Larissa Rapadas"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Improving analysis code by writing functions

The function `prot_analysis ()` below takes any PDB protein structure and prints a plot of the protein's B-factor over the length of its residues.

```{r}
# First, we have to load the bio3d package into our workspace
# The pdb functions will not work without it
library (bio3d)

prot_analysis <- function (pdb) { # This takes any PDB acc. # in "_"
 
  # Read the PDB file
  protein <- read.pdb (pdb)
  
  # Trim PBD to only atoms of chain A and C-alpha
  chainA <- trim.pdb (protein, chain="A", elety="CA")
  
  # Select the B-factor data from the chainA subset above
  protein.b <- chainA$atom$b
  
  # Plot the B-factor data along the residues of the protein input
  # Will also denote the corresponding secondary structure along the margin
  plotb3 (protein.b, sse=chainA, typ="l", ylab="B-factor") 
}
```

Now, we can call this function for any protein of our choosing,
as long as it has a PDB accessible file.

Let's try calling `prot_analysis ()` on the one of the kinases we tried in class.
```{r}
prot_analysis ("4AKE")
```

When called correctly, the function should output a plot of the protein's B-factor 
over the length of its residues. There should also be marginal rectangles corresponding 
to the protein's secondary structure at that specific residue: gray = sheets, black = helices.

                                                      -- END --
